### 2018-02-01 04-related.tex
- Primeiro commit - Apenas compila - Semântica antiga (Tese e artigo TECS) - Texto novo (SBLP)

\section{Related Work}

\CEU was strongly influenced by Esterel but they differ in the most fundamental 
aspect of the notion of time~\cite{ceu.sensys13}.
%
Esterel defines time as a discrete sequence of logical unit instants or 
``ticks''.
At each tick, the program reacts to an arbitrary number of simultaneous input 
events from the environment.
%
The presence of multiple inputs requires careful static analysis to detect and
reject programs with \emph{causality cycles} and \emph{schizophrenia problems}%
~\cite{esterel.constructive,esterel.d7,esterel.d6,esterel.d3,esterel.d5,esterel.d8,esterel.d1,esterel.schizo2}.
%
In contrast, \CEU defines time as a discrete sequence of reactions to
unique input events.
%
In the formal semantics, ... \fs{como isso aparece na semantica} ...
\CEU also rejects some syntactically correct programs to avoid infinite
execution, but with simple restrictions in the abstract syntax tree.

Another distinction is that, in Esterel, the behavior of internal and external
events is equivalent, while in \CEU internal events introduce stack-based micro
reactions within external reactions, providing more fine-grained control for
intra-reaction execution.
%
In the formal semantics, ... \fs{como isso aparece na semantica} ...
%
Some variants of the Statecharts synchronous visual language also distinguish
internal from external events~\cite{statecharts.variants}.
In Statemate~\cite{statecharts.statemate}, \emph{``reactions to external and
internal events (...) can be sensed only after completion of the step''},
implying queue-based execution.
In Stateflow~\cite{statecharts.stateflow}, \emph{``the receiving state (of the
event) acts here as a function''}, which is similar to \CEU's
stack-based execution.
We are not aware of formalizations for these ideas for a deeper comparison with
\CEU.

Like other synchronous languages
(%
    \emph{Reactive C}~\cite{rp.rc},                 % 1991
    \emph{Protothreads}~\cite{wsn.protothreads},    % 2006
    \emph{SOL}~\cite{wsn.sol},                      % 2007
    \emph{SC}~\cite{rp.synchc},                     % 2009
    and
    \emph{PRET-C}~\cite{rp.pretc}%                  % 2010
),
\CEU relies on deterministic scheduling to preserve intra-reaction determinism.
%
In addition, it also performs concurrency checks to detect trails that, when
reordered, change the observable behavior of the program, i.e., trails that
actually rely on deterministic scheduling~\cite{ceu.sensys13}.
%
Esterel is only deterministic with respect to external behavior: \emph{``the
same sequence of inputs always produces the same sequence of 
outputs''}~\cite{esterel.primer}.
%
However, the execution order for operations within a reaction is 
non-deterministic: \emph{``if there is no control dependency, as in
\code{(call~f1()~||~call~f2())},
the order is unspecified and it would be an error to rely on 
it''}~\cite{esterel.primer}.
%
For this reason, Esterel, does not support shared-memory concurrency:
\emph{``if a variable is written by some thread, then it can neither be read
nor be written by concurrent threads''}~\cite{esterel.primer}.

Esterel describes a finalization mechanism in a standardization
proposal~\cite{esterel.v7} that is similar to \CEU's.
However, we are not aware of an open implementation or a formal semantics for a
deeper comparison.

\fs{outras linguagens sincronas}
\fs{outras linguagens deterministicas}
\fs{outras linguagens com terminacao}