### 2018-02-01 index.tex

\documentclass[sigplan, anonymous, review]{acmart}

\input{header}
% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
  Paso, Texas USA}
\acmYear{1997}
\copyrightyear{2016}

\acmPrice{15.00}

%\acmBadgeL[http://ctuning.org/ae/ppopp2016.html]{ae-logo}
%\acmBadgeR[http://ctuning.org/ae/ppopp2016.html]{ae-logo}

\begin{document}

\title[A Deterministic and Terminating Semantics for \CEU]
{A Deterministic and Terminating Semantics
  for the Synchronous Programming Language~\CEU}

\author[R.\,C.\,M.~Santos]{Rodrigo C.\,M.~Santos}
\email{rsantos@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}
%
\author[G.\,F.~Lima]{Guilherme F.~Lima}
\email{glima@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}
%
\author[F.~Sant'Anna]{Francisco Sant'Anna}
\email{francisco@ime.uerj.br}
\affiliation{UERJ, Rio de Janeiro, Brazil}
%
\author[R.~Ierusalimschy]{Roberto Ierusalimschy}
\email{roberto@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}
%
\author[E.\,H.~Haeusler]{Edward Hermann Haeusler}
\email{hermann@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}

\begin{abstract}
\CEU is a synchronous programming language for embedded soft real-time systems.
%
It supports advanced features, such as safe shared-memory concurrency and
abortion of execution lines, while enforcing internal determinism and
time-bounded reactions.
%
In this paper, we present a small-step structural operational semantics for
\CEU and prove that under it program reactions are deterministic and always
terminate.
%
Thus, given the same timeline of input events, the same \CEU program in the
same initial memory state will always react in bounded time and arrive at the
same final memory state.
%
We also discuss some equivalence results and the relation between the proposed
semantics and its actual implementation.
\end{abstract}

\terms{Languages, design, theory}
\keywords{Operational semantics, \CEU, Synchronous languages}

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10003752.10010124.10010131.10010134</concept_id>
  <concept_desc>Theory of computation~Operational semantics</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10011007.10011006.10011008.10011009.10011014</concept_id>
  <concept_desc>Software and its
                engineering~Concurrent programming languages</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010562.10010564</concept_id>
  <concept_desc>Computer systems
                organization~Embedded software</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}
\ccsdesc[500]{Theory of computation~Operational semantics}
\ccsdesc[500]{Software and its engineering~Concurrent programming languages}
\ccsdesc[300]{Computer systems organization~Embedded software}

\maketitle

\input{01-intro}
\input{02-ceu}
\input{03-semantics}
\input{04-related}
\input{05-conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{my,other}

\appendix
%\include{proofs}

\end{document}