### 2018-02-22 05-conclusion.tex

\section{Conclusion}
\label{sec.conclusion}

The programming language \CEU aims to offer a concurrent, safe, and realistic
alternative to C for embedded soft real-time systems, such as sensor networks
and multimedia systems.
%
%Its synchronous semantics enables safe shared-memory concurrency and
%safe abortion of lines of execution, while enforcing memory-bounded,
%deterministic, and terminating reactions to the environment.
%
\CEU inherits the synchronous and imperative mindset of Esterel but adopts a
simpler semantics with fine-grained execution control, which makes the
language fully deterministic.
%
In addition, its stack-based execution for internal events provides a limited
but memory-bounded form of subroutines.
%
\CEU also provides a finalization mechanism for resources when interacting with
the external environment.

We propose a small-step structural operational semantics for \CEU and a proof
that reactions are deterministic, terminate in finite time, and use bounded
memory, i.e., that for a given arbitrary timeline of input events, multiple
executions of the same program always react in bounded time and arrive at the
same final finite memory state.