### 2018-03-19 index.tex
Typos levantados pelos revisores

\documentclass[sigplan,anonymous,review,protrusion=true,expansion]{acmart}

\input{header}
% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[LCTES'18]{ACM SIGPLAN/SIGBED}{June 2018}{Philadelphia, USA}
%\acmYear{1997}
%\copyrightyear{2016}

%\acmPrice{15.00}

%\acmBadgeL[http://ctuning.org/ae/ppopp2016.html]{ae-logo}
%\acmBadgeR[http://ctuning.org/ae/ppopp2016.html]{ae-logo}

\begin{document}

\title[A Memory-Bounded, Deterministic and Terminating Semantics for~\CEU]
{A Memory-Bounded, Deterministic and Terminating Semantics for the
       Synchronous Programming Language~\CEU}

\author[R.\,C.\,M.~Santos]{Rodrigo C.\,M.~Santos}
\email{rsantos@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}
%
\author[G.\,F.~Lima]{Guilherme F.~Lima}
\email{glima@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}
%
\author[F.~Sant'Anna]{Francisco Sant'Anna}
\email{francisco@ime.uerj.br}
\affiliation{UERJ, Rio de Janeiro, Brazil}
%
\author[R.~Ierusalimschy]{Roberto Ierusalimschy}
\email{roberto@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}
%
\author[E.\,H.~Haeusler]{Edward Hermann Haeusler}
\email{hermann@inf.puc-rio.br}
\affiliation{PUC-Rio, Rio de Janeiro, Brazil}

\begin{abstract}
\CEU is a synchronous programming language for embedded soft real-time systems.
%
It focuses on control-flow safety features, such as safe shared-memory
concurrency and safe abortion of lines of execution, while enforcing
memory-bounded, deterministic, and terminating reactions to the environment.
%
In this work, we present a small-step structural operational semantics for
\CEU and a proof that reactions have the properties enumerated above:
%
that for a given arbitrary timeline of input events, multiple executions of the
same program always react in bounded time and arrive at the same final finite
memory state.
%
%We also discuss some equivalence results and the relation between the proposed
%semantics and its actual implementation.
\end{abstract}

\terms{Languages, design, theory}
\keywords{Determinism, Termination, Operational semantics, Synchronous languages}

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10003752.10010124.10010131.10010134</concept_id>
  <concept_desc>Theory of computation~Operational semantics</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10011007.10011006.10011008.10011009.10011014</concept_id>
  <concept_desc>Software and its
                engineering~Concurrent programming languages</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010562.10010564</concept_id>
  <concept_desc>Computer systems
                organization~Embedded software</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}
\ccsdesc[500]{Theory of computation~Operational semantics}
\ccsdesc[500]{Software and its engineering~Concurrent programming languages}
\ccsdesc[300]{Computer systems organization~Embedded software}

\maketitle

\input{01-intro}
\input{02-ceu}
\input{03-semantics}
\input{04-related}
\input{05-conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{my,other}

\input{06-appendix}

\end{document}