### 2018-05-03 07-artifact.tex
Improve Artifact Evaluation based on the reviews.

\section{Artifact Appendix}

%Submission and reviewing guidelines and methodology:
%{\em http://cTuning.org/ae/submission.html}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Abstract}

Our artifact includes an open-source implementation of the programming language
\CEU.
The implementation is based on and should conform with the formal semantics
presented in this paper.
The artifact also includes an executable script with over 3500 test cases of
valid and invalid programs in \CEU.
The script is customizable and allows to create new tests providing inputs and
expected outputs.
The evaluation platform is a Linux/Intel with Lua-5.3 and GCC-7.2 installed.

\subsection{Artifact check-list (meta-information)}

{\small
\begin{itemize}
  %\item {\bf Algorithm: }
  %\item {\bf Program: }
  \item {\bf Compilation: } The output of the \CEU compiler is a C program that requires a C compiler (e.g., GCC-7.2).
  \item {\bf Transformations: } The compiler of \CEU is a Lua program (Lua-5.3) that generates a C program.
  %\item {\bf Binary: }
  \item {\bf Data set: } The data set is a set of program test cases included in the language distribution.
  \item {\bf Run-time environment: } Linux with Lua-5.3 and GCC-7.2. Root access is required to install a single executable file.
  \item {\bf Hardware: } An off-the-shelf Intel machine.
  %\item {\bf Run-time state: }
  \item {\bf Execution: } 5-10 minutes for the full test.
  \item {\bf Output: } Console output: Success (termination) or Fail (abortion).
  \item {\bf Experiments: } Manual steps performed by the user in the command line.
  %\item {\bf Workflow frameworks used?: }
  %\item {\bf Publicly available?: } Yes: \url{https://github.com/fsantanna/ceu}
\end{itemize}

\begin{itemize}
  \item {\bf Artifacts publicly available?:} Yes.
  \item {\bf Artifacts functional?:} Yes.
  \item {\bf Artifacts reusable?:} No.
  \item {\bf Results validated?:} No.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description}

\subsubsection{How delivered}

The compiler/distribution of \CEU is available on GitHub:

\url{https://github.com/fsantanna/ceu}

The website of \CEU includes an introductory video and an online tutorial:

\url{http://www.ceu-lang.org/}

The language also provides extensive documentation online:

\url{http://fsantanna.github.io/ceu/out/manual/v0.30/}

\subsubsection{Hardware dependencies}

An off-the-shelf Intel machine.

\subsubsection{Software dependencies}

Linux, Lua-5.3 (with lpeg-1.0.0), and GCC-7.2.

\subsubsection{Data sets}

A set of more than 3500 programs packed in a single script file which is
included with the compiler distribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Installation}

Install all required software (assuming an Ubuntu-based distribution):

\begin{verbatim}
$ sudo apt-get install git lua5.3 lua-lpeg liblua5.3-0 \
                          liblua5.3-dev
\end{verbatim}

Clone the repository of \CEU:

\begin{verbatim}
$ git clone https://github.com/fsantanna/ceu
$ cd ceu/
$ git checkout v0.30
\end{verbatim}

Install \CEU:

\begin{verbatim}
$ make
$ sudo make install     # install as "/usr/local/bin/ceu"
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment workflow}

Run the experiment:

\begin{verbatim}
$ cd tst/
$ ./run.lua
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and expected result}

The experiment will execute all test cases.
In about 5-10 minutes, a summary will be printed on screen:

\begin{verbatim}
$ cd tst/
$ ./run.lua
<...> # output with the test programs
stats = {
    count = 3478,
    trails = 9082,
    bytes = 50803896,
    bcasts = 0,
    visits = 5189770,
}
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment customization}

\subsubsection{Test Cases}

The file \code{tst/tests.lua} includes all test cases.
Each test case contains a program in \CEU as well as the expected result, e.g.:

\begin{verbatim}
Test { [[
    var int ret = 0;
    var int i;
    loop i in [0 -> 10[ do
        ret = ret + 1;
    end
    escape ret;
]],
    run = 10,
}
\end{verbatim}

This test case should compile and run successfully yielding \code{10}.

To customize the experiment, include a new test case at line \code{440}, after
the string ``\code{--~OK:~well~tested}''.

\subsubsection{Sample Programs}

The distribution of \CEU comes with sample programs that can be executed as
follows:

\begin{verbatim}
$ cd /<...>/ceu/               # change to the Céu repository
$ make samples                  # execute all samples one by one
\end{verbatim}

\subsubsection{Single File}

It is also possible to compile a single file with a program in \CEU:

\begin{verbatim}
$ cd /<...>/ceu/                         # change to the Céu repository
$ make one CEU_SRC=/tmp/tst.ceu # compile a program
$ /tmp/tst                      # execute it
\end{verbatim}

As an example, the file \code{/tmp/tst.ceu} may contain a program such as
follows:

\begin{verbatim}
input int A; // two input events
input int B;

spawn do     // generates test inputs asynchronously
    await async do
        emit A(10);
        emit B(100);
    end
end

var int a = await A; // main program
var int b = await B;

escape a+b;  // 110 is the result of the program back to the OS
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Notes}